---
title: "ps1_Puqiu"
author: "Puqiu Zhao"
format: html
editor: visual
---

```{r}
#| message: false 
library(tidyverse)
```

## Part 1

Build the table

```{r}
unit <- c(1:5) 
Y_i1 <- c(8, 6, 3, 5, 2) 
Y_i0 <- c(4, 5, 3, 7, 6) 
D_i <- c(1, 1, 0, 0, 0) 
table_1 <- tibble(unit, Y_i1, Y_i0, D_i)
table_1
```

### Q1: Compute $\tau_i$ for each unit and the $\textbf{ATE}$

```{r}
table_1 <- table_1 |> mutate(tau_i = Y_i1 - Y_i0)
table_1
table_1 |> 
  summarise(ATE = mean(tau_i))
```

### Q2: Compute $\textbf{ATT}$ and $\textbf{ATC}$

```{r}
#ATT
table_1 |> 
  filter(D_i == 1) |> 
  summarise(ATT = mean(tau_i))

#ATC
table_1 |> 
  filter(D_i == 0) |> 
  summarise(ATC = mean(tau_i))
```

### Q3: Using switching equation, compute observed $DiM$

```{r}
#switching equation
table_1 <- table_1 |> 
  mutate(Y_obs = D_i * Y_i1 + (1- D_i) * Y_i0)

#DiM
table_1 |> 
  summarise(DiM = mean(Y_obs[D_i == 1]) - mean(Y_obs[D_i == 0]))
```

### Q4: Is DiM too big? Too small? Identify selection bias

The $DiM$ is $1.67$, which is too big. The $ATE$ is $-0.2$. Therefore, there is a selection bias.

## Part 2

### Q1: DAG graph

$$ D \leftarrow M \rightarrow Y $$ $$ D \rightarrow Y $$

The backdoor path is $D \leftarrow M$ and $M \rightarrow Y$

### Q2: confounder

### Q3: What should you condition on to satisfy backdoor criterion?

To satisfy the backdoor criterion, we should condition on $M$. Conditioning on $M$ can block the backdoor path $D \leftarrow M \rightarrow Y$.

### Q4: If $M$ is unobserved, what are your options?

We can conduct a randomized controlled trial (RCT) to eliminate the confounding effect of prior motivation ($M$). By randomly assigning individuals to the job training program, the treatment becomes independent of prior motivation, therefore blocking the backdoor path.

## Part 3

### Q1:

```{r}
patients <- tibble(
id = 1:10,
y1 = c(7, 5, 5, 7, 4, 10, 1, 5, 3, 9), # Y(1)
y0 = c(1, 6, 1, 8, 2, 1, 10, 6, 7, 8) # Y(0)
)
# We know ATE = 0.6 from the perfect doctor
#Randomly assign 7 to surgery, 3 to chemo — 10,000 times:
set.seed(0818)
n_reps <- 10000
dim_vec <- numeric(n_reps)

for (r in 1:n_reps){
  d <- sample(c(rep(1, 7), rep(0, 3))) #7 treated (surgery), 3 untreated (chemo)
  y_obs <- d*patients$y1 + (1-d)*patients$y0 #switching equation
  dim_vec[r] <- mean(y_obs[d == 1]) - mean(y_obs[d == 0]) #DiM
}

mean(dim_vec)
sd(dim_vec)
```

Yes, $DiM$ is still approximately equal to the $ATE$.

### Q2:

```{r}
patients_1 <- patients |> 
  select(-y0) |> #remove old y0
  mutate(y0_man = y1 - 0.6) #manually y0

set.seed(0818)
n_reps <- 10000
dim_vec_man <- numeric(n_reps)

for (r in 1:n_reps){
  d <- sample(c(rep(1, 7), rep(0, 3)))
  y_obs <- d*patients_1$y1 + (1-d)*patients_1$y0_man #switching equation
  dim_vec_man[r] <- mean(y_obs[d == 1]) - mean(y_obs[d == 0]) #DiM
}

mean(dim_vec_man)
sd(dim_vec_man)
```

We can compare the distributions of the DiM estimator before and after manually modifying the control group. Although the two distributions have similar means, their standard deviations differ. In particular, the distribution of the DiM estimator becomes more dispersed after the control group is manually altered.

### Q3:

```{r}
d_bias <- as.integer(rank(-patients$y1) <= 5) #"-"5 largest values
y_obs_bias <- d_bias*patients$y1 + (1-d_bias)*patients$y0
dim_bias <- mean(y_obs_bias[d_bias == 1]) - mean(y_obs_bias[d_bias == 0])

mean(dim_bias)
```
